---
title: "Legacy and next-generation data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Legacy and next-generation data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(zipcodeR)
```

## Which interface should I use?

ZIP and spatial data change over time, but silently replacing package data can
change already-published analyses. zipcodeR therefore separates the historical
contract from corrected, versioned behavior.

For a new analysis, use the `_ng` functions with an explicitly selected modern
bundle. This is the forward-looking interface documented for new community
code. Calling an `_ng` function communicates that the analysis opted into the
bundle's newer data and semantics; it does **not** opt into automatic future
updates. Keep the version and SHA pinned for the life of the project.

Use the unsuffixed functions when maintaining, rerunning, or comparing existing
zipcodeR code. They are the historical compatibility interface.

All names that existed in 0.3.5 keep the same data, algorithm, ordering,
rounding, warnings, errors, and unusual edge cases. This is intentional even
where a behavior would be designed differently today.

```{r}
zip_distance("08731", "08901")
get_cd("08731")
reverse_zipcode(c("08731", "08999", "08731"))
zip_data_version()
```

The returned metadata identifies the legacy snapshot. No option or network
state can switch these functions to newer data.

## Start a new analysis by pinning a modern bundle

A data release contains one RDS bundle and a JSON manifest. The manifest lists
the bundle SHA256, every raw source URL and SHA256, licenses, vintages, build
commit and R version, dependency-lock checksum, schemas, row counts, and
canonical output hashes.

After a version has been published and registered, download it explicitly:

```{r, eval = FALSE}
bundle <- download_zip_data_bundle("2026.08")
```

For offline or archival work, keep the RDS file with the project and verify the
manifest checksum while reading it:

```{r, eval = FALSE}
bundle <- read_zip_data_bundle(
  "data/zipcodeR-data-2026.08.rds",
  sha256 = "SHA256_FROM_THE_RELEASE_MANIFEST"
)
```

Aliases such as `latest` are rejected. A lookup never downloads a missing
bundle or falls back to a different version.

## Use the `_ng` interface

Every data-dependent next-generation function takes the bundle first:

```{r, eval = FALSE}
reverse_zipcode_ng(bundle, c("08731", "08999", "08731"))
geocode_zip_ng(bundle, c("08731", "08999", "08731"))
search_radius_ng(bundle, 39.9, -74.3, radius = 10)
zip_distance_ng(bundle, "08731", "08901")
get_tracts_ng(bundle, "08731")
get_cd_ng(bundle, "08731")
```

These functions preserve input order and duplicates, make missing records
explicit, validate inputs consistently, keep geographic identifiers as
characters, and use the bundle's declared authoritative mappings and vintage.
USPS-only ZIPs without an authoritative district relationship remain unmapped
with a quality reason.

## Recording provenance in research outputs

Store the complete version record, not merely the human-readable version:

```{r, eval = FALSE}
version_record <- zip_data_version(bundle)
saveRDS(version_record, "results/zipcodeR-data-version.rds")

version_record$data_version
version_record$bundle_sha256
zip_data_provenance(bundle, dataset = "zip_to_cd", key = "08731")
```

For a reproducibility supplement, retain the release manifest, the bundle, and
the analysis code together. The verified `bundle_sha256` attribute is added by
`read_zip_data_bundle()` and `download_zip_data_bundle()` and is propagated to
`_ng` lookup results.
