---
title: "Frequently Asked Questions"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Frequently Asked Questions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(zipcodeR)
```

## Why are the longitude values negative?

Because the United States is in the western hemisphere. Geographic
coordinates follow a standard sign convention: latitudes north of the
equator are positive, and longitudes **west of the prime meridian are
negative**. Every location in the U.S. therefore has a negative
longitude:

```{r}
geocode_zip("90210")
```

This is not an error, and you should not multiply `lng` by -1: mapping
libraries, spatial packages, and web APIs all expect this convention.

## Why is a valid ZIP code missing from `zip_code_db`?

Two common reasons:

1. **Data vintage.** The bundled database is a snapshot; ZIP codes
   introduced after the snapshot date will be absent until the next data
   refresh. Check `zip_code_db` directly to see whether a code is
   present.
2. **ZIP codes are not ZCTAs.** USPS ZIP codes are collections of
   delivery routes and points, not areas. The Census Bureau's ZIP Code
   Tabulation Areas (ZCTAs) approximate most — but not all — ZIP codes.
   Some USPS-only codes (typically P.O. Box or single-building "unique"
   codes, such as 91230 in Glendale, CA) have no ZCTA and no
   Census-derived attributes, and may be missing from ZCTA-oriented
   sources. Use `is_zcta()` to check whether a given ZIP code is also a
   ZCTA.

Relatedly, about 20% of the ZIP codes in `zip_code_db` (P.O. Box and
unique codes) have no `lat`/`lng` coordinates; functions such as
`zip_distance()` return `NA` distances for them.

## How accurate is the city / county ("jurisdiction") information?

Treat it as approximate. A ZIP code is a postal-delivery construct, not
a municipal boundary: one ZIP code can cross city, county, and even
state lines, and the USPS "city" name is the preferred *mailing* name,
which frequently differs from the municipality a given address actually
lies in (addresses just outside a city's limits often carry that city's
name). The `county` column reflects the predominant county for the ZIP
code. If you need authoritative jurisdiction assignment, geocode the
full street address against boundary files (e.g., Census TIGER/Line)
rather than relying on the ZIP code alone.

## I get `object 'zip_code_db' not found`

The datasets are lazy-loaded with the package. This error almost always
indicates a broken or partial installation — for example, a package
library that was migrated between R versions. Reinstalling usually
fixes it:

```r
install.packages("zipcodeR")
```

If you want the data without attaching the package, use
`zipcodeR::zip_code_db`.

## Where does the data come from?

`zip_code_db` derives from the
[uszipcode](https://github.com/MacHu-GWU/uszipcode-project) project's
database, which aggregates U.S. Census Bureau data (decennial census,
ACS, gazetteer files) with USPS-derived ZIP code attributes.
`zcta_crosswalk` comes from the Census Bureau's ZCTA-to-tract
relationship files, and `zip_to_cd` from HUD-USPS crosswalk data. See
the help page of each dataset for details and known limitations.
